---
title: "Workshop 1"
output: 
  flexdashboard::flex_dashboard:
    storyboard: true
    theme: cosmo
date: "2024-02-14"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Corpus

When do we call a pop song an indiepop song? Last month, the famous 2000’s pop song ‘Murder on the Dancefloor’ by Sophie Ellis-Bextor has been rediscovered by TikTok after the movie Saltburn came out. The song was covered by ‘indie-pop’ duo Royel Otis. After the cover the song turned into a timeless indiepop song with sounds from the 2010’s. As a musician, myself heavily inspired by the indie-pop style I am more than interested to discover what musical components make a regular pop song differ from an indie-pop song that makes me and many others feel a certain nostalgic way. My corpus is genre-based but I plan to analyze artists, songs, playlists, and charts with this research question. I hope to be able to discover when Spotify recognizes songs within Spotify as ‘indie-pop’, ‘pop’, or maybe in some cases both. I am not sure how deep we can analyze the song's components (chords, scales etc.) or just the metadata around it. I plan to get the tracks from playlists as “Indie Pop”, “Indie Pop Hits” and “Pop Rising” curated by Spotify. As specific tracks I want to look at current “pop” hits such as “Flowers” by Miley Cyrus and “Greedy” by Tate McRae and current “indie pop” hits such as the Murder On The Dancefloor cover by Royel Otis and a classic as Kilby Girl by The Backseat Lovers.



### Plot

```{r echo=FALSE}
library(tidyverse)
library(spotifyr)
library(plotly)
library(flexdashboard)

indiepop <- get_playlist_audio_features("", "37i9dQZF1DWWEcRhUVtL8n?si=8549c3d965e4458c&nd=1&dlsi=7e635eb660464a36")

pop <- get_playlist_audio_features("", "37i9dQZF1DWXti3N4Wp5xy?si=e0a5344cb1b545c5&nd=1&dlsi=61886b9565ee4c74")

music <-
  bind_rows(
    indiepop |> mutate(category = "Indie Pop Playlist"),
    pop |> mutate(category = "Pop Playlist")
  )

#music |>
 # ggplot(aes(x = energy)) +
  #geom_histogram(binwidth = 0.1) +
  #facet_wrap(~category)


ggplot <- music %>%
  ggplot(aes(x = energy, y = acousticness)) +  # Specify y variable as well
  geom_point(aes(color = category), alpha = 0.7) +  # Use geom_point for scatterplot, add color aesthetics for category
  facet_wrap(~category) +  # Faceting by category
  labs(x = "Energy", y = "Acousticness") +  # Adding axis labels
  theme_minimal()  # Ap

ggplotly(ggplot)

```

### Analysis




```{r }
```
